# Reproducible Research: Peer Assessment 1

Set the global options

```{r global_options, echo=TRUE}
opts_chunk$set(fig.width=12, fig.height=8, fig.align='center', fig.path='figures/',
               echo=TRUE, warning=FALSE, message=FALSE)
```

## Loading and preprocessing the data

```{r reading the data}
if (!file.exists("activity.csv")) {
    unzip("activity.zip")
}

AMdata = read.csv("activity.csv", na.strings = "NA")
```



## What is mean total number of steps taken per day?

```{r }

steps_perday = aggregate(AMdata$steps, by=list(AMdata$date), FUN=sum, na.rm=TRUE)

hist(steps_perday[,2],xlab="steps",breaks=10,main="Histogram of steps",col="red")
```

The **mean** total number of steps taken per day is

```{r}
mean(steps_perday[,2])
```

The **median** total number of steps taken per day is

```{r}
median(steps_perday[,2])
```


## What is the average daily activity pattern?

```{r }

steps_interval = aggregate(AMdata$steps,by=list(AMdata$interval),FUN=function(x) mean(x,na.rm=TRUE))

plot(steps_interval,type="l",xlab="Interval",ylab="average steps",main="average steps per interval",col="blue")

```


The `r steps_interval[steps_interval[,2] == max(steps_interval[,2]),][[1]]` 5-minute interval ,which is the `r as.numeric(row.names(steps_interval[steps_interval[,2] == max(steps_interval[,2]),]))`th row, contains the maximum number of steps.

```{r}
steps_interval[steps_interval[,2] == max(steps_interval[,2]),]
```



## Imputing missing values

```{r }
na_row = nrow(AMdata[is.na(AMdata$steps),])

AMdata_new=AMdata

set.seed(100)

AMdata_new[is.na(AMdata_new$steps),]$steps = sample(steps_interval[,2],na_row,replace = TRUE)

steps_perday_new=aggregate(AMdata_new$steps,by=list(AMdata_new$date),FUN=function(x) sum(x,na.rm=TRUE))

```

1. The total number of missing values in the dataset is `r na_row`.

2. The strategy adopted here is sampling from the the mean for that 5-minute interval. 

3. The new dataset without `NA` is `AMdata_new`.

4. The histogram of the new dataset

```{r}
hist(steps_perday_new[,2],xlab="steps",breaks=10,main="Histogram of steps",col="red")
```

The **mean** total number of steps taken per day of the new dataset is

```{r}
mean(steps_perday_new[,2])
```

The **median** total number of steps taken per day of the new dataset is

```{r}
median(steps_perday_new[,2])
```

```{r}
library(knitr)

d = data.frame(mean_value=c(mean(steps_perday[,2]),mean(steps_perday_new[,2])),median_value=c(median(steps_perday[,2]),median(steps_perday_new[,2])),row.names=c("old_data","new_data"))

kable(d)
```


## Are there differences in activity patterns between weekdays and weekends?

Set the language

```{r }
Sys.setlocale("LC_TIME", "C")
```

`dayind` is a new factor variable in the dataset with two levels ¨C ¡°weekday¡± and ¡°weekend¡± indicating whether a given date is a weekday or weekend day.


```{r}
date_st = as.Date(AMdata_new$date,format = "%Y-%m-%d")

dayind = {weekdays(date_st,abbreviate = TRUE)=="Sun" | weekdays(date_st,abbreviate = TRUE)=="Sat"}

dayind[dayind == TRUE] = "weekend"
dayind[dayind == FALSE] = "weekday"
dayind = as.factor(dayind)
```

Make a panel plot

```{r}
steps_interval_weekday = with(AMdata_new[dayind=="weekday",],aggregate(steps,by=list(interval),FUN="mean"))

steps_interval_weekday$dayind = "weekday"

steps_interval_weekend = with(AMdata_new[dayind=="weekend",],aggregate(steps,by=list(interval),FUN="mean"))

steps_interval_weekend$dayind = "weekend"

newdata = rbind(steps_interval_weekday, steps_interval_weekend)

colnames(newdata) = c("interval","steps_avg","dayind")

library(lattice)

xyplot(steps_avg~interval | dayind, data=newdata , type="l",layout=c(1,2),ylab="Number of steps",xlab="Interval")

```


